---
title: "Sinclair Conglomerate Analysis"
author: "Joseph E. Panzik"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
```

## Overview
This data is used and published in Panzik, J.E., Evans, D.A.D., Kasbohm, J.J., Hanson, R.E., Gose, W. & Desormeau, J. Using paleomagnetism and geochronologic data to determine late Mesoproterozoic paleogeographic history and tectonic relations of the Sinclair Terrane, Namaqua orogen, Namibia. *Geological Society of London Special Publication* **424**, 119-143 (2016).

The analysis below shows and describes the conglomerate test performed for the publication and provides open-access to the data.

## Loading Data
The data will be loaded from a CSV file that contains the site information and least squares declination and inclination from demagnetization for each sample.

```{r}
cong <- read.csv(file="data/Sinclair 2011-2012 Conglomerates.csv", header=TRUE, na.strings=NA)
```

The data

```{r}
str(cong)
head(cong)
```


```{r}
p<-ggplot(cong, aes(x=Temp)) + 
      geom_histogram(aes(y=..density..), color="black", fill="white") +
      geom_density()
p
```