---
title: "Sinclair Conglomerate Analysis"
author: "Joseph E. Panzik"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(kableExtra)
opts_chunk$set(echo = TRUE)
```

## Overview
This data is used and published in Panzik, J.E., Evans, D.A.D., Kasbohm, J.J., Hanson, R.E., Gose, W. & Desormeau, J. Using paleomagnetism and geochronologic data to determine late Mesoproterozoic paleogeographic history and tectonic relations of the Sinclair Terrane, Namaqua orogen, Namibia. *Geological Society of London Special Publication* **424**, 119-143 (2016).

The analysis below shows and describes the conglomerate test performed for the publication and provides open-access to the data.

## Loading Data
The data will be loaded from a CSV file that contains the site information and least squares declination and inclination from demagnetization for each sample.

```{r}
cong <- read.csv(file="data/Sinclair 2011-2012 Conglomerates.csv", header=TRUE, na.strings=NA)
```

## Basic Structures & Plots of Conglomerate Data
The data

```{r}
str(cong)
head(cong)
```

The first thing I want to see is the distribution of demagnetization temperature related to the demagnetization of magnetite at ~580$^\circ$C and hematite at ~685$^\circ$C.
```{r}
ggplot(cong, aes(x=Temp)) + 
  geom_histogram(aes(y=..density..), color="black", fill="white", bins=20) +
  geom_density() +
  geom_vline(aes(xintercept=580), color="blue") +
  geom_vline(aes(xintercept=685), color="red") +
  xlab("Demagnetization Temperature (°C)") +
  ggtitle("Distribution of Demagnetization Temperatures")
```

## Data Filtering

Some of the samples have multiple temperature entries, so the lower temperature magnetic direction will be ignored in favor of the higher temperature directionality that should be consistent with hematite.

```{r}
cong <- cong[!duplicated(cong[ , c("Site","Sample")], fromLast=TRUE),]
```

## Determining Randomness

Visual inspectio  of randomness is done through plotting the inclination and declination values on stereonets.

Since JP1137 & JP1214 came from the same location, and JP1145 & JP1209 came from the same location, individual and combined locality stereonets will be created.

**Note: The $a_{95}$ circles are not perfect, but do a good job estimating the uncertainty.**

The r code chunk for JP1137 is shown below, but the code chunks for other stereonets will be hidden from the final document to save space.

```{r, out.width="150%", out.height="150%"}
#Radius scaling factor
f<-1.36
#Create JP1137 data subset
sub <- cong[cong$Site=="JP1137",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1137 Conglomerate Magnetic Directions")
p
```

```{r, echo=FALSE, out.width="150%", out.height="150%"}
#Create JP1214 data subset
sub <- cong[cong$Site=="JP1214",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1214 Conglomerate Magnetic Directions")
p
#Create JP1137 & JP1214 data subset
sub <- cong[cong$Site=="JP1137"|cong$Site=="JP1214",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1137 & JP1214 Conglomerate Magnetic Directions")
p
#Create JP1145 data subset
sub <- cong[cong$Site=="JP1145",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1145 Conglomerate Magnetic Directions")
p
#Create JP1209 data subset
sub <- cong[cong$Site=="JP1209",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1209 Conglomerate Magnetic Directions")
p
#Create JP1145 & JP1209 data subset
sub <- cong[cong$Site=="JP1145"|cong$Site=="JP1209",]
p <- ggplot(sub) +
  geom_hline(yintercept = seq(0, 90, by = 30), colour = "grey90", size = 0.5) +
  geom_vline(xintercept = seq(0, 360-1, by = 45), colour = "grey90", size = 0.5) +
  coord_polar() +
  scale_x_continuous(limits=c(0,360), breaks=seq(0,315,by=45)) +
  scale_y_reverse(limits=c(90,0), breaks=seq(90,0,by=-30)) +
  scale_radius(range=c(f*min(sub$a95),f*max(sub$a95))) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.key = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        panel.grid  = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(sub$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  ggtitle("JP1145 & JP1209 Conglomerate Magnetic Directions")
p
```

The function of a conglomerate test is to determine if the clasts have a magnetic overprint that has affected the directionality of them all. While the image above indicates there appears to be randomness, we will compare the visual randomness to that of Watson's (1956) test for randomness.

Magnetic directions must first be converted to Cartesian coordinates.

```{r}
rad <- pi/180
cong$x <- cos(cong$Inc*rad)*cos(cong$Dec*rad)
cong$y <- cos(cong$Inc*rad)*sin(cong$Dec*rad)
cong$z <- sin(cong$Inc*rad)
```

The Watson (1956) randomness test compares the vector length created by summing the Cartesian coordinates, $R=\sqrt{(x_{sum})^2+(y_{sum})^2+(z_{sum})^2}$, to a critical value, $R_0=\sqrt{7.815*\frac{n}{3}}$, within each site. The $R$ value can range from $0$ to $n$ the number of samples used in the site; the closer the value is to $0$, the more random it is.

Since JP1137 & JP1214 came from the same location, and JP1145 & JP1209 came from the same location, $R$ and $R_0$ values will be determined for each individual site and for the site pairs.

Sample code chunk is shown for JP1137. Code chunks for other sites are not displayed in output for brevity.

```{r}
#Site JP1137
xSumJP1137 <- sum(cong$x[cong$Site=="JP1137"])
ySumJP1137 <- sum(cong$y[cong$Site=="JP1137"])
zSumJP1137 <- sum(cong$z[cong$Site=="JP1137"])
RJP1137 <- sqrt((xSumJP1137)^2+(ySumJP1137)^2+(zSumJP1137)^2)
R0JP1137 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1137"])/3))
```

```{r, echo=FALSE}
#Site JP1214
xSumJP1214 <- sum(cong$x[cong$Site=="JP1214"])
ySumJP1214 <- sum(cong$y[cong$Site=="JP1214"])
zSumJP1214 <- sum(cong$z[cong$Site=="JP1214"])
RJP1214 <- sqrt((xSumJP1214)^2+(ySumJP1214)^2+(zSumJP1214)^2)
R0JP1214 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1214"])/3))
#Combined JP1137 & JP1214
xSumJP1137JP1214 <- xSumJP1137+xSumJP1214
ySumJP1137JP1214 <- ySumJP1137+ySumJP1214
zSumJP1137JP1214 <- zSumJP1137+zSumJP1214
RJP1137JP1214 <- sqrt((xSumJP1137JP1214)^2+(ySumJP1137JP1214)^2+(zSumJP1137JP1214)^2)
R0JP1137JP1214 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1137"|cong$Site=="JP1214"])/3))
#Site JP1145
xSumJP1145 <- sum(cong$x[cong$Site=="JP1145"])
ySumJP1145 <- sum(cong$y[cong$Site=="JP1145"])
zSumJP1145 <- sum(cong$z[cong$Site=="JP1145"])
RJP1145 <- sqrt((xSumJP1145)^2+(ySumJP1145)^2+(zSumJP1145)^2)
R0JP1145 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1145"])/3))
#Site JP1209
xSumJP1209 <- sum(cong$x[cong$Site=="JP1209"])
ySumJP1209 <- sum(cong$y[cong$Site=="JP1209"])
zSumJP1209 <- sum(cong$z[cong$Site=="JP1209"])
RJP1209 <- sqrt((xSumJP1209)^2+(ySumJP1209)^2+(zSumJP1209)^2)
R0JP1209 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1209"])/3))
#Combined JP1145 & JP1209
xSumJP1145JP1209 <- xSumJP1145+xSumJP1209
ySumJP1145JP1209 <- ySumJP1145+ySumJP1209
zSumJP1145JP1209 <- zSumJP1145+zSumJP1209
RJP1145JP1209 <- sqrt((xSumJP1145JP1209)^2+(ySumJP1145JP1209)^2+(zSumJP1145JP1209)^2)
R0JP1145JP1209 <- sqrt(7.815*(length(cong$x[cong$Site=="JP1145"|cong$Site=="JP1209"])/3))
```
We can now put the $R$ and $R_0$ values for each in a table to determine whether sites pass the randomness test. The $R_{max}$ value (maximum possible length of $R$) is also shown for each site as a comparison. Sites that fail the Watson (1956) randomness test because $R>R_0$ will be highlighted in light red.

```{r}
R <- c(RJP1137, RJP1214, RJP1137JP1214, RJP1145, RJP1209, RJP1145JP1209)
R_0 <- c(R0JP1137, R0JP1214, R0JP1137JP1214, R0JP1145, R0JP1209, R0JP1145JP1209)
R_max <- c(length(cong$x[cong$Site=="JP1137"]),length(cong$x[cong$Site=="JP1214"]),length(cong$x[cong$Site=="JP1137"|cong$Site=="JP1214"]), length(cong$x[cong$Site=="JP1145"]), length(cong$x[cong$Site=="JP1209"]), length(cong$x[cong$Site=="JP1145"|cong$Site=="JP1209"]))
RTest <- data.frame(R, R_0, R_max, row.names=c("JP1137", "JP1214", "JP1137 & JP1214", "JP1145", "JP1209", "JP1145 & JP1209"))
RTest %>%
  kbl() %>%
  kable_styling(full_width = F, html_font = "Cambria",
                position = "left", bootstrap_options ="hover") %>%
  column_spec(2, color=ifelse(RTest$R>RTest$R_0, "white", "black"),
              background=ifelse(RTest$R>RTest$R_0, "lightcoral", "white")) %>%
  column_spec(1, bold=TRUE)
```

## Discussion
This shows that overall the conglomerate sites pass the Watson (1956) randomness test with site $R$ values being smaller than site $R_0$ values. The exception being **JP1214** and **combined JP1137 & JP1214**. There is a cluster within JP1214 that show a remanace of north and steeply-downward. The origin of the overprint is unknown