---
title: "Sinclair Strat Sections"
author: "Joseph E. Panzik"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(sqldf)
opts_chunk$set(echo = TRUE)
```

## Overview

This analysis goes through samples and sites collected in the Sinclair Terrane that are stratigraphic sedimentary sections. Bedding corrections are applied to the samples in 10% unfolding increments with an applied K Test.

Pole positions are determined for each sample and section after bedding corrections are applied.

## Loading Data

The data is loaded from a CSV file that contains the site information and least squares declination and inclination from demagnetization for each sample.

```{r}
strat <- read.csv(file="data/Sinclair 2011-2012 Strat.csv", header=TRUE, na.strings=NA)
```

## Basic Structures & Plots
The data

```{r}
str(strat)
head(strat)
```

## Unfolding Analysis

A fold test is performed to determine if there is a magnetic overprint since deformation. This is done by applying a 10% incremental unfolding of each sample and performing a K Test within each site and on all of the sites combined.

The right-hand strike of bedding was done without an adjustment for present magnetic field declination variation. This means that a site variation adjustment will be applied to the right-hand strike based on the site coordinates (latitude & longitude) determined by the [NOAA Magnetic Declination Calculator](https://www.ngdc.noaa.gov/geomag/calculators/magcalc.shtml).

```{r}
strat$RHS_Adjust <- strat$Bed_RHS+strat$Site_Var
```

```{r}
#rad <- pi/180
#c <- abs(90-cong$Lat)
#a <- 90-atan(0.5*tan(cong$Inc*rad))/rad
#b <- acos(cos(a*rad)*cos(c*rad)+sin(a*rad)*sin(c*rad)*cos(cong$Dec*rad))/rad
#A <- asin(sin(a*rad)*sin(cong$Dec*rad)/sin(b*rad))/rad

#Pole_Lat <- 90-b
#Pole_Long <- cong$Long+A

#Pole_Lat
#Pole_Long
```