---
title: "Sinclair Conglomerate Analysis"
author: "Joseph E. Panzik"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(gcookbook)
opts_chunk$set(echo = TRUE)
```

## Overview
This data is used and published in Panzik, J.E., Evans, D.A.D., Kasbohm, J.J., Hanson, R.E., Gose, W. & Desormeau, J. Using paleomagnetism and geochronologic data to determine late Mesoproterozoic paleogeographic history and tectonic relations of the Sinclair Terrane, Namaqua orogen, Namibia. *Geological Society of London Special Publication* **424**, 119-143 (2016).

The analysis below shows and describes the conglomerate test performed for the publication and provides open-access to the data.

## Loading Data
The data will be loaded from a CSV file that contains the site information and least squares declination and inclination from demagnetization for each sample.

```{r}
cong <- read.csv(file="data/Sinclair 2011-2012 Conglomerates.csv", header=TRUE, na.strings=NA)
```

The data

```{r}
str(cong)
head(cong)
```

The first thing I want to see is the distribution of demagnetization temperature related to the demagnetization of magnetite at ~580$^\circ$C and hematite at ~680$^\circ$C.
```{r}
ggplot(cong, aes(x=Temp)) + 
  geom_histogram(aes(y=..density..), color="black", fill="white", bins=20) +
  geom_density() +
  geom_vline(aes(xintercept=580), color="blue") +
  geom_vline(aes(xintercept=680), color="red") +
  xlab("Demagnetization Temperature (°C)") +
  ggtitle("Distribution of Demagnetization Temperatures")
```

```{r}
p <- ggplot(cong) +
  coord_polar() +
  geom_point(aes(x=Dec, y=abs(Inc), size=a95), fill=ifelse(cong$Inc>=0,"red","#00BFC4"), shape=21, alpha=0.3) + 
  geom_point(aes(x=Dec, y=abs(Inc), shape=Inc>=0)) +
  scale_shape_discrete(name = "Inclination", labels = c("Negative", "Positive")) +
  scale_size_area(max_size=10) +
  scale_x_continuous(limits=c(0,360),breaks=seq(0,360,by=45)) +
  xlab("Declination (°)") +
  ylab("Inclination (°)") +
  ggtitle("Magnetic Directions of Conglomerate Samples")
p
```